Tracing the Meta-Level: PyPy's Tracing JIT Compiler

### 2009-06-09 talk/icooolps2009/paper.tex
Changes (thanks Anto).

There are quite an number of approaches that try to minimally enhance
interpreters to generate code at runtime without actually writing a native
compiler by hand. The goal of these is to get rid of dispatch overhead of
typical interpreters while retaining ease of implementation. Piumarta and
Riccardi \cite{piumarta_optimizing_1998} propose to copy fragments of the
interpreter together for commonly occurring bytecode sequences to reduce
dispatch overhead. However, dispatching is still needed to jump between such
sequences and also when non-copyable bytecodes occur. Ertl and Gregg
\cite{ertl_retargeting_2004} go further and stitch together the concatenated
sequences by patching the copied machine code. Thus they get rid of all dispatch
overhead. Both techniques can speed up interpreters which large dispatch
overhead a lot. However they will help less if the bytecodes themselves do a
lot of work (as is the case with Python), whereas our tracing technique can
trace into the implementation of those bytecodes.

>>>>

There are quite a number of approaches that try to minimally enhance
interpreters to generate code at runtime without actually writing a native
compiler by hand. The goal of these is to get rid of dispatch overhead of
typical interpreters while retaining ease of implementation. Piumarta and
Riccardi \cite{piumarta_optimizing_1998} propose to copy fragments of the
interpreter together for commonly occurring bytecode sequences to reduce
dispatch overhead. However, dispatching is still needed to jump between such
sequences and also when non-copyable bytecodes occur. Ertl and Gregg
\cite{ertl_retargeting_2004} go further and stitch together the concatenated
sequences by patching the copied machine code. Thus they get rid of all dispatch
overhead. Both techniques can speed up interpreters which large dispatch
overhead a lot. However they will help less if the bytecodes themselves do a
lot of work (as is the case with Python) and the dispatch overhead is lower. On
the other hand, our technique can do a better job by tracing inside the
implementation of those bytecode and inlining common paths.
